Understanding fine temporal scale evolutionary dynamics with lineage metrics

### 2018-04-02 paper.tex
Update on Overleaf.

@@

\title{Something about lineage metrics}
%\author{Chris Adami$^{1}$, Ryoichi Seki$^{1,2}$ \and Robel Yirdaw$^2$ \\
%\mbox{}\\
%$^1$California Institute of Technology, Pasadena, CA 91125 \\
%$^2$California State University, Northridge, CA 91330 \\
%adami@caltech.edu} % email of corresponding author

@@

Evolution is an emergent effect of events (\textit{e.g.} replication, variation, and competition) that occur on a much finer-grained temporal scale. While this is a large part of why it is so fascinating, it also presents challenges to studying the short-term mechanisms that govern long-term results. In computational evolutionary systems, we theoretically have adequate data to untangle these mechanisms. In practice, though, gathering appropriate data to do so can be challenging. Moreover, because of the vast difference between the scales that the causes and effects occur on, the quantity of such data can be overwhelming. Both of these problems can be mitigated with a standardized set of diagnostic summary metrics that paint a picture of shorter-term evolutionary dynamics within a population. Here, we introduce a suite of such metrics that operate on lineages and provide an intuition for what they can tell us about evolution.

@@

A lineage describes a continuous line of descent, linking ancestors and their descendants. A complete lineage can provide a post-hoc, step-by-step guide to the evolution of an extant organism where each step involves replication and inherited variation. Indeed, lineage analyses are a powerful tool for disentangling the evolution of complexity in both natural and digital systems; digital systems, however, allow for fine-grained lineage tracking at resolutions that are impossible in modern wet lab experiments, allowing us to replay the tape of life in exacting detail to tease apart the evolutionary recipe for whatever complex feature we happen to be interested in. 
Yet, tracking the full details of a single lineage, much less a population of lineages, can be computationally expensive and will inevitably generate an unwieldy amount of data that can be challenging to visualize and interpret [CITATIONS]. Summary statistics can help alleviate these issues by focusing on general trends(?) for a population rather than for each individual.
% @AML: Need a transition from here to next paragraph about how diagnostic/summary statistics can help
% @SPJ: Something like this?

Here, we demonstrate three types of lineage measurements: lineage length, mutation accumulation, and phenotypic volatility. For each, we discuss their application and our intuitions for what they can tell us about evolution. We evaluate our intuitions on a set of two-dimensional, real-valued optimization problems under a range of mutation rates and selection strengths.   
% @AML: we'll want to say something here about how we talk about how these metrics extend to genetically discrete organisms (e.g. GP)

\section{Metrics}
% Outlining some points to (maybe) make here; probably too bogged down in details; not even sure we want to talk about lineages as graphs:
% - Apply metrics in context of asexual populations
% - Establish a framework from which to think about metrics (not novel, but rarely formalized)
% 	 - We treat lineages as sequences of states where each state represents an individual and transitions between states represent parent-offspring relationships.
%      - More formally, lineages are directed acyclic graphs where internal nodes have in-degree = 1 (parent), out-degree = 1 (offspring). Original ancestor (in digital evolution, often seed of population) w/out-degree = 1. Extant individual w/in-degree = 1.
%      - Transitions specify relevant variation introduced into offspring (e.g. mutations)
%      - States specify relevant genotype and phenotype information (e.g. genome, location in space, behavior, etc) -- anything we might want to measure over a lineage
%      - Becomes more complicated with sexual populations, but we'll stick with asexual assumption for this. 
%  - Already, framework affords certain questions:
%     - How many individuals are along the lineage? 
%     - Total # and magnitude of mutations along lineage?
%     - Distribution of mutation types (ins vs. sub/beneficial vs. deleterious)?
%  - Further abstractions:
%      - Sequence of genotypes along lineage (where each genotype may represent multiple individuals)
%      - Sequence of phenotypes along lineage (where each phenotype may represent multiple individuals)
%      - Still further abstraction: sequence of a particular aspect of genotype (e.g. genetic marker)/phenotype (e.g. location) along lineage
%  - We discuss three classes of lineage metrics within this framework: lineage length, mutation accumulation, and phenotypic volatility. 
%  - For each, we discuss our intuition for what they can tell us (alone and in combination) about evolution.

\subsection{Lineage Length}
% Define/explanation
% 	- Within context of framework, lineage length is simply measure of number of states in the graph/sequence.
%   - What this means will depend on what states represent?
%       - number of individuals along lineage 
%             - More interesting in experiments where generations are not dictated by system, but rather by life history strategies of organisms
%             - Give you a diagnostic for replication rate (r vs. k strategy)
%      - number of genotypes along lineage
%            - length = 1: no mutations
%            - length = number of individuals: mutation in every replication
%            - Intuitively, should be higher at high mutation rates, lower at low mutation rates. 
%      - number of phenotypes along lineage
%            - low: evidence for static environment
%            - high: might be indicative of a changing environment or ecological interactions
%            - if number of genotypes is high, but number of phenotypes low, probably in a neutral fitness landscape
%      - various subsets of genotype/phenotype
%           - location in space (environment type): useful for looking at lineage through heterogeneous landscape
%           - change in particular phenotypic trait
%           - change in genetic markers
%           - Distributions of above two lengths can tell you useful things about drift, selection, etc.

@@

\subsection{Mutation Accumulation}
Mutation Accumulation defines a set of measurements that track mutational changes across a lineage. These changes can be measured in a variety of ways, such as the magnitude of the change (for real-valued genomes), or the total count of changes (for discrete-valued genomes). The type of a mutation can also be tracked, and can provide insight into the distribution of mutation types for a given lineage. In conjunction with collected fitness information, the class of a mutation (\textit{e.g.} beneficial, deleterious, or neutral) can also be tracked. % @SPJ: TODO
% Mutation accumulation over a lineage 
%  - Measured in a variety of ways
%   - For real-valued genomes, we can track magnitudes of changes
%   - For discrete-valued genomes, we can track counts of changes
%      - Track by type to get distribution of mutation types
%   - In conjunction with fitness information, can track classes of mutation: beneficial, deleterious, neutral
%      - Again, can extract a distribution; different environments/experiments, you might expect different distributions
%      - Importance of deleterious steps; valley crossing, etc

\subsection{Phenotypic Volatility}
% Define/describe
% .... 
% Useful things that this metric could identify:
% - Mutational bet hedging (high volatility)
% - Plasticity (high volatility)
% - Neutral landscapes (expect low volatility)

\subsection{Summary statistics}

@@

@@

\section{Data Analysis}

% Trim dramatically
%Evolutionary biologists often accuse each other of making up "just-so" stories - post-hoc explanations for how an observed trait or pattern could have evolved. The problem with these stories is that, in biology, there is often no way to verify that they reflect what happened in biology. Due to evolution's stochastic nature, it is easy to come up with possible ways that something could have happened once, regardless of how repeatable they are. Theoretically, we should easily be able to avoid this problem in computational evolution by carefully checking the underlying assumptions behind our hypothesized explanation.  All too often, however, we fail to drill down to the true underlying mechanism behind an observed effect.

@@

\section{Results and Discussion}
% Changing mutation: here's what happens when we change mutation on each problem
\begin{figure}
\includegraphics[width=3.5in]{figs/dom_deleterious_mutation_rate.png}
\end{figure}
\begin{figure}
\includegraphics[width=3.5in]{figs/variance_deleterious_mutation_rate.png}
\end{figure}

@@

>>>>

% *****************
%  TODOs:
% *****************
% [ ] Writing: Abstract
% [ ] Writing: Results & discussion
% [ ] Writing: Conclusions
% [ ] Big 'ol citation pass: need to work appropriate citations into paper

@@

\title{Understanding fine temporal scale evolutionary dynamics with lineage metrics}
\author{Emily Dolson$^{1,2,3}$, Alexander Lalejini$^{1,2, 3}$, Steven Jorgensen$^{1,2}$ \and Charles Ofria$^{1,2, 3}$ \\
\mbox{}\\
$^1$BEACON Center for the Study of Evolution in Action, Michigan State University, East Lansing, MI, 48824 \\
$^2$Department of Computer Science and Engineering, Michigan State University, East Lansing, MI, 48824 \\
$^3$Ecology, Evolutionary Biology, and Behavior Program, Michigan State University, East Lansing, MI, 48824 \\
dolsonem@msu.edu} % email of corresponding author

@@

Evolution is an emergent effect of events (\textit{e.g.} replication, variation, and competition) that occur on a much finer-grained temporal scale. While evolution's emergent nature is a large part of why it is so fascinating, it also presents challenges to studying the short-term mechanisms that govern long-term results. In computational evolutionary systems, we theoretically have adequate data to untangle these mechanisms. In practice, though, gathering appropriate data to do so can be challenging. Moreover, because of the vast difference between the scales that evolutionary causes and effects can occur on, the quantity of such data can be overwhelming. Both of these problems can be mitigated with a standardized set of diagnostic summary metrics that paint a picture of shorter-term evolutionary dynamics within a population. Here, we suggest a suite of such metrics that operate on lineages and phylogenies. We then use experimental results to provide an intuition for what they can tell us about evolution.

@@

A lineage describes a continuous line of descent, linking ancestors and their descendants. A complete lineage can provide a post-hoc, step-by-step guide to the evolution of an extant organism where each step involves replication and inherited variation. Indeed, lineage analyses are a powerful tool for disentangling evolutionary dynamics in both natural and digital systems; digital systems, however, allow for fine-grained lineage tracking at resolutions that are impossible in modern wet lab experiments, allowing us to replay the tape of life in exacting detail and to tease apart the evolutionary recipe for whatever complex feature or phenomenon we happen to be interested in. 
Yet, tracking the full details of a single lineage, much less a population of lineages, can be computationally expensive and will inevitably generate an unwieldy amount of data that can be challenging to visualize and interpret [CITATIONS].
% @AML: Need a transition from here to next paragraph about how diagnostic/summary statistics can help
% @SPJ: Something like this?
%@ELD That looks like a good start, but we still need the part where we  actually discuss our methodology for creating these metrics. I'm going to try splitting it off into its own paragraph

Summary statistics can help alleviate these issues by enabling the user to focus on aggregate trends across a population rather than needing to examine each individual. The question is how to effectively summarize what is essentially a path through fitness space. A useful abstraction is to treat the path as a sequence of states. Here, we will use phenotypes and genotypes as the states in the sequence, but we could just as easily some other descriptor of the lineage's position in the fitness landscape at a given point in time. % @ELD: Should we cite Alex's phenotypic plasticity stuff here? @AML: Nah, this paragraph is pretty sleek in making its points. I think the metrics section might be more appropriate. We can use it as an example. 
With this abstraction in hand, a few metrics are easily formalized: the number of unique states, the number of transitions between states, and the amount of time spent in each state. 
Additionally, we may care about how the transitions between states happened. What mutations lead to them? Were those mutations beneficial, deleterious, or neutral at the time? Here, we choose a subset of these metrics that we expect will be particularly useful to explore further.

Whereas a lineage represents the evolutionary history that lead up to a single individual, a phylogeny represents the evolutionary history of an entire population. Measurements that summarize phylogenies can provide useful insight into population-level evolutionary dynamics, such as diversification and co-existence among different clades. A variety of useful phylogeny measurements have already been developed by biologists~\citep{tucker_guide_2017}. These measurements tend to treat the phylogeny as a graph and make calculations about its topology. Tucker \textit{et al.} group them into three broad categories: assessments of the quantity of evolutionary history represented by a population, assessments of the amount of divergence within that evolutionary history, and assessments of the topological regularity of the phylogenetic tree. Such measurements can help quantify the behavior of the population as a whole, providing insight into interactions between its members. Thus, they are useful indicators of the presence of various types of eco-evolutionary dynamics.
% @AML: I think the next bit (just moved it) might belong in the metrics section. Instead, we might want a little more elaboration on each of the categories of metrics.

Here, we present three types of lineage measurements: lineage length, mutation accumulation, and phenotypic volatility. Additionally, we present four phylogeny measurements: most-recent common ancestor depth, phylogenetic richness, phylogenetic divergence, and phylogenetic regularity. For each metric, we discuss its application and our intuitions for what it can tell us about evolution. We evaluate our intuitions on a set of two-dimensional, real-valued optimization problems under a range of mutation rates and selection strengths. For this work, we restrict our attention to asexually reproducing populations; however, we make suggestions as to how these metrics can be extended to sexual populations.

\section{Metrics}

\subsection{Lineage Metrics}
% Outlining some points to (maybe) make here; probably too bogged down in details; not even sure we want to talk about lineages as graphs:
% - Apply metrics in context of asexual populations
% - Establish a framework from which to think about metrics (not novel, but rarely formalized)
% 	 - We treat lineages as sequences of states where each state represents an individual and transitions between states represent parent-offspring relationships.
%      - More formally, lineages are directed acyclic graphs where internal nodes have in-degree = 1 (parent), out-degree = 1 (offspring). Original ancestor (in digital evolution, often seed of population) w/out-degree = 1. Extant individual w/in-degree = 1.
%      - Transitions specify relevant variation introduced into offspring (e.g. mutations)
%      - States specify relevant genotype and phenotype information (e.g. genome, location in space, behavior, etc) -- anything we might want to measure over a lineage
%      - Becomes more complicated with sexual populations, but we'll stick with asexual assumption for this. 
%  - Already, framework affords certain questions:
%     - How many individuals are along the lineage? 
%     - Total # and magnitude of mutations along lineage?
%     - Distribution of mutation types (ins vs. sub/beneficial vs. deleterious)?
%  - Further abstractions:
%      - Sequence of genotypes along lineage (where each genotype may represent multiple individuals)
%      - Sequence of phenotypes along lineage (where each phenotype may represent multiple individuals)
%      - Still further abstraction: sequence of a particular aspect of genotype (e.g. genetic marker)/phenotype (e.g. location) along lineage
%  - We discuss three classes of lineage metrics within this framework: lineage length, mutation accumulation, and phenotypic volatility. 
%  - For each, we discuss our intuition for what they can tell us (alone and in combination) about evolution. 
Each of the three lineage metrics we discuss --- lineage length, mutation accumulation, and phenotypic volatility --- reduces a lineage to a linear sequence of states where each state represents an individual or a continuous sequence of individuals that share some common genotypic or phenotypic characteristic of interest. While we limit our focus to three lineage metrics, this abstraction places lineages in a form suitable for a wide range of measurements, including the direct application of many data mining techniques designed to operate over sequences. % @AML: TODO - will want to come up with a few examples/citations for such data mining techniques (can refer to data mining class notes/textbook).

% @AML: this while paragraph could use some sanity checking/smoothing out. 
Only lineages consisting of asexually reproducing ancestors where genetic material is exclusively vertically transmitted can be directly abstracted as a \textit{linear sequence} of states. 
Sexual reproduction (and any form of horizontal gene transfer) complicates matters significantly as such lineages are more appropriately represented by trees rooted at the extant organism, branching for each contributor of genetic material. 
For each metric we present here, we limit our discussion to asexual populations; however, we suggest two approaches for generalizing to sexual populations: apply a lossy compression to reduce sexual lineages into linear sequences or extend our application of the metrics to operate over non-linear (tree) structured lineages. 
Sexual lineages can be compressed into linear sequences of states by modeling sexual reproduction events as asexual reproduction events, designating one parent to be a part of the lineage and considering the genetic contributions of other parents as sources of genetic variation (mutations). As previously stated, this approach is lossy, discarding potentially important parentage information. 
Alternatively, we can extend our metrics to operate over the more complex state sequences that constitute the lineages of sexually reproducing organisms. One such approach would be to consider all possible ancestor paths for an extant individual, taking an average of a measurement over each path.

\subsubsection{Lineage Length}
% Define/explanation
% 	- Within context of framework, lineage length is simply measure of number of states in the graph/sequence.
%   - What this means will depend on what states represent?
%       - number of individuals along lineage 
%             - More interesting in experiments where generations are not dictated by system, but rather by life history strategies of organisms
%             - Give you a diagnostic for replication rate (r vs. k strategy)
%      - number of genotypes along lineage
%            - length = 1: no mutations
%            - length = number of individuals: mutation in every replication
%            - Intuitively, should be higher at high mutation rates, lower at low mutation rates. 
%      - number of phenotypes along lineage
%            - low: evidence for static environment
%            - high: might be indicative of a changing environment or ecological interactions
%            - if number of genotypes is high, but number of phenotypes low, probably in a neutral fitness landscape
%      - various subsets of genotype/phenotype
%           - location in space (environment type): useful for looking at lineage through heterogeneous landscape
%           - change in particular phenotypic trait
%           - change in genetic markers
%           - Distributions of above two lengths can tell you useful things about drift, selection, etc.

Lineage length describes the number of ancestral states traversed by a lineage. If a state represents a single individual along a lineage, lineage length is a measure of the number of generations on a lineage. This form of the lineage length metric is most useful in systems where generational turnover is not algorithmically determined but instead determined by the life history strategies of organisms. In these systems, lineage length can provide a proxy for replication rate where, for lineages that span equal lengths of time, longer lineages imply faster replication rates (\textit{e.g.} an r replicator strategy) and shorter lineages imply slower replication rates (\textit{e.g.} a k replicator strategy).

Lineage length becomes an increasingly flexible and informative metric if we begin to consider more abstract definitions of an ancestral state along a lineage. For example, we might measure the length of a lineage according to a particular phenotypic trait or genetic marker where an ancestral state represents a temporally continuous sequence of individuals that share the characteristics of interest.

\subsubsection{Mutation Accumulation}
Mutation Accumulation defines a set of measurements that track mutational changes across a lineage. These changes can be measured in a variety of ways, such as the magnitude of the change (for real-valued genomes), or the total count of changes (for discrete-valued genomes). The type of a mutation can also be tracked, and can provide insight into the distribution of mutation types for a given lineage. Measures of mutation accumulation along the lineages of successful extant individuals can help to tease apart the relative importance of different types of mutational events when compared to what is expected by chance.

In conjunction with collected fitness information, the class of a mutation (\textit{e.g.} beneficial, deleterious, or neutral) can also be tracked. Different evolutionary conditions are expected to cause different distributions of mutations along a lineage [CITE barrick and lenski genome dynamics]; deviations revealed by measures of mutation accumulation can act as a barometer for unexpected evolutionary dynamics.   
The number and magnitude of deleterious mutations along a lineage can both tell us something about the ruggedness of the fitness landscape as well as how well a lineage was able to cross fitness valleys. 
An elevated measure of neutral mutations relative to beneficial or deleterious mutations might indicate that a lineage is spending most of its time drifting in a neutral fitness landscape.

% Mutation accumulation over a lineage 
%  - Measured in a variety of ways
%   - For real-valued genomes, we can track magnitudes of changes
%   - For discrete-valued genomes, we can track counts of changes
%      - Track by type to get distribution of mutation types
%   - In conjunction with fitness information, can track classes of mutation: beneficial, deleterious, neutral
%      - Again, can extract a distribution; different environments/experiments, you might expect different distributions
%      - Importance of deleterious steps; valley crossing, etc

\subsubsection{Phenotypic Volatility}

Phenotypic volatility addresses the rate at which the lineage changes phenotype. In systems with discrete/categorical phenotypes, this can be measured by summing the number of times the phenotype changes. A related but subtly different measurement in such systems is the number of unique phenotypes on a lineage. In most cases, these values will be very similar. A discrepancy between them, however, would be a clear signal of interesting dynamics. Specifically, it would suggest that the lineage were in a portion of the landscape where it could rapidly switch among different phenotypes. Such behavior could be indicative of some form of evolutionary bet-hedging.

In systems with continuous-valued phenotypes, a subtly different approach is needed to measure phenotypic volatility. 
% TODO

% Define/describe
% .... 
% Useful things that this metric could identify:
% - Mutational bet hedging (high volatility)
% - Plasticity (high volatility)
% - Neutral landscapes (expect low volatility)

@@

\subsubsection{Summary statistics}

@@

A single representative lineage can be chosen by selecting the lineage of the dominant organism (either the most fit or the more numerous; here we use the most fit). In populations with a diversity of coexisting strategies this approach can be uninformative as the dominant lineage is not necessarily representative of all successful lineages. An alternative in these scenarios is to filter out lineages that do not have offspring some predetermined number of generations later as such lineages were likely not representative of an important subset of the population. Still, any approach based on measuring only a subset of lineages can be challenging to interpret when the current dominant lineage (or lineages) is replaced with a different one; such changes can introduce a discontinuity if the value is being measured over time. If graceful responses to changes in which lineage is dominant are required, it can be advantageous to instead measure summary statistics (e.g. mean, variance, and range) across the entire population.

In scenarios where selective sweeps happen rapidly, the dominant lineage will likely be very similar to the average lineage in the population, as most of the population will be closely related. When the population contains more phylogenetic diversity, however, it is possible for the dominant lineage to differ from the mean. The nature of such differences is likely informative about the evolutionary dynamics occurring in the population. Since we have yet to observe an instance of this phenomenon, we do not explore it further in this paper. Nevertheless, it presents an argument for measuring the dominant lineage in addition to gathering summary statistics across the population.

\subsection{Phylogeny metrics}

These metrics operate on entire phylogenies rather than single lineages within a population. Thus, in most cases, summarizing them across the population is not an issue as it is with lineage metrics. Because they factor in the entire population, phylogeny metrics can be more computationally expensive to calculate than metrics about a single lineage. On the other hand, because most lineages in the population share a substantial amount of history, phylogeny metrics can often be calculated more rapidly than lineage metrics can be calculated across an entire population. Note that phylogenies can be constructed with regard to any taxonomic level of organization, be it individual, genotype, phenotype, etc. Thus, when we refer to items in a phylogeny, we will use the term \textit{taxa}.

A standard technique for saving memory and time when working with lineages in computational systems is to "prune" them, removing any taxa that are not either currently alive or the ancestor of taxa that are alive. Since all of the phylogeny metrics we discuss here are borrowed from natural systems (where we do not have information about taxa without offspring), they all are designed to work on pruned phylogenies. Thus, for the remainder of this paper, we will assume we are only working with pruned phylogenies.

In populations without ecological forces promoting coexistence, phylogenies should coalesce periodically, resulting in pruned lineages that mostly consist of a single path. When there is strong selection, this coalescence should happen even more rapidly. Thus, phylogenies with topologies that deviate from that expectation are an indication of ecological interactions within the population. The metrics discussed here can provide insight into the nature of those interactions and their long-term evolutionary effects. As a result, they are often referred to as phylogenetic diversity metrics~\citep{tucker_guide_2017}.

An important distinction between phylogenies in natural vs computational systems to be aware of is that in natural systems phylogenies are generally inferred from extant taxa, whereas in computational systems they are directly measured. The inferred phylogenies used in natural systems do not contain internal nodes unless those nodes are branch points. They also do not contain history prior to the most recent common ancestor (MRCA) of all organisms in the extant population. For consistency, we will exclude pre-MRCA taxa from our analyses. However, we will not remove non-branching internal nodes, as these only serve to make our phylogenies more informative.

Here we provide a high-level summary of phylogeny metrics that we expect will be particularly useful. For more metrics and more detail on all of these metrics, see~\citep{winter_phylogenetic_2013, tucker_guide_2017}.

\subsubsection{Depth of Most-Recent Common Ancestor}

The depth of the MRCA (i.e. how many steps from the beginning of time did it occur?) is a useful metric because of the relatively high ratio of informativeness to ease of calculation. A more recent MRCA implies more frequent selective sweeps and less long-term stable coexistence between clades. The depth of the MRCA is so easily calculated because most any implementation of phylogeny pruning requires keeping track of it already. Measuring the frequency with which the MRCA changes (i.e. the number of coalescence events) can also by informative, as some conditions can inflate the length of the lineage relative to other conditions, without actually increasing the frequency of selective sweeps. This scenario is particularly likely when the population size is changing over time. A downside to the depth of MRCA as a metric is that any population that does have a stable ecology will likely never see a change in the MRCA after the very beginning of evolution.

\subsubsection{Phylogenetic Richness}

Measurements of phylogenetic richness quantify the total amount of evolutionary history that is contained in a set of taxa. The most traditional metric of phylogenetic richness is the one simply titled "Phylogenetic Diversity", which is calculated as the number of the nodes in the minimum spanning tree from the MRCA to all extant taxa~\citep{faith_conservation_1992}. Another approach is to calculate the pairwise distances between all taxa and sum them~\citep{tucker_guide_2017}. A third approach is to calculate evolutionary distinctiveness, a measurement of a taxon's evolutionary uniqueness~\citep{isaac_mammals_2007}, and sum it across all extant taxa~\citep{tucker_guide_2017}.

\subsubsection{Phylogenetic Divergence}

Measurements of phylogenetic divergence quantify how distinct the taxa in the population are from each other, and are often the means of values that can be calculated for individual taxa. For instance, one option is to calculate the mean of the pairwise distances between all taxa in the population~\citep{webb_exploring_2000}. Similarly, phylogenetic divergence can also be calculated by calculating the evolutionary distinctiveness of each taxon in the population (as alluded to in the previous section) and taking the average of these values.

\subsubsection{Phylogenetic Regularity}

Measurements of phylogenetic regularity quantify how balanced the branches of the phylogeny are, and are often the variances of values that can be calculated for individual taxa. Just as the mean of the pairwise distances between all taxa in the population is a measurement of phylogenetic divergence, taking their variance is a measurement of phylogenetic regularity. The same is true of the variance of evolutionary distinctiveness across the population.

@@

\section{Data Analysis}
\subsection{3D visualizations}
% Trim dramatically
%Evolutionary biologists often accuse each other of making up "just-so" stories - post-hoc explanations for how an observed trait or pattern could have evolved. The problem with these stories is that, in biology, there is often no way to verify that they reflect what happened in biology. Due to evolution's stochastic nature, it is easy to come up with possible ways that something could have happened once, regardless of how repeatable they are. Theoretically, we should easily be able to avoid this problem in computational evolution by carefully checking the underlying assumptions behind our hypothesized explanation.  All too often, however, we fail to drill down to the true underlying mechanism behind an observed effect.

@@

\subsection{Metric analysis}

We analyzed trends in our metrics using the R Statistical Computing Language~\citep{r_core_team_r:_2017}. Specifically, we used the ggplot2 library for all graphs included in this paper~\citep{wickham_ggplot2:_2009}. All analysis scripts are available at https://github.com/stevenjson/ALife2018-Lineage.

\section{Results and Discussion}
% Changing mutation: here's what happens when we change mutation on each problem
\begin{figure}
\includegraphics[width=3.5in]{figs/mean_deleterious_mutation.png}
\end{figure}
\begin{figure}
\includegraphics[width=3.5in]{figs/mrca_depth_mutation.png}
\end{figure}

% \begin{figure}
% \includegraphics[width=3.5in]{figs/variance_deleterious_mutation_rate.png}
% \end{figure}

@@

\begin{figure}
\includegraphics[width=3.5in]{figs/mean_deleterious_ts.png}
\end{figure}
\begin{figure}
\includegraphics[width=3.5in]{figs/mrca_ts.png}
\end{figure}
